---
title: 'Plotting NPS Data'
subtitle: 'A Step-by-Step Walkthrough for Plotting NPS Data using GGPlot'
author: 'Author: [Chris Mahoney](https://www.linkedin.com/in/chrimaho/)'
date: 'Published: TBA'
#date: 'Date: `r format(Sys.Date(), "%d/%b/%Y")`'
toc-title: Contents
output:
  html_document:
    code_download: yes
    highlight: haddock
    number_sections: yes
    template: default_toc.html
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
    includes:
      in_header: header.html
      after_body: footer.html
  html_notebook:
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: no
    includes:
      in_header: header.html
      after_body: footer.html
---

<style>
.math {
    <!-- font-size: 120%; -->
    font-style: normal;
    font-family: "Cambria Math";
}
.column {
    float: left;
    width: 50%;
    border: 1px solid black;
}
.row:after {
    content: "";
    display: table;
    clear: both;
}
h1, .h1 {
    margin-top: 40px;
    font-weight: bold;
}
h2, .h2 {
    margin-top: 40px;
    margin-left: 40px;
}
</style>


<!-- Acknowledgements: -->



<!-- Set Up Environment -->

```{r SET Function, echo=FALSE, eval=TRUE}
# Define function to load packages ----
LoadPackages <- function(packages, install=FALSE) {
    
    # Input:
    # - 'packages' : An atomic string or a string vector of the list of packages to load.
    # - 'install'  : A boolean value for whether or not to install the packages that are missing.
    
    # Output:
    # - A logical result (TRUE of FALSE) for if they were successfully loaded.
    
    # Validations:
    stopifnot(is.character(packages))
    stopifnot(is.logical(install))
    
    # Remove all packages. Note: The suppression functions are to limit the amount of printed output.
    for (package in .packages()) {
        if (!package %in% c("parallel", "stats", "graphics", "grDevices", "datasets", "utils", "methods", "base")) { #THESE PACKAGES ARE PART OF BASE!! YOU CANNOT REMOVE THEM!! But you can remove everything else...
            suppressPackageStartupMessages ( 
                suppressMessages ( 
                    suppressWarnings ( 
                        detach ( paste0("package:", package) #The `detach()` function is like the reverse of `library()` or `require()`.
                                 , unload = TRUE
                                 , character.only = TRUE
                        )
                    )
                )
            )
        }
    }
    
    # Install all defined packages
    if (install==TRUE) {
        for (package in packages) {
            if (!package %in% installed.packages()) { #The `installed.packages()` function  returns a vector of all the installed packages...
                install.packages ( package
                                   # , quiet = TRUE
                                   # , verbose = FALSE
                                   , dependencies = TRUE
                )
            }
        }
    }
    
    # Load all defined packages
    for (package in packages) { #Need to loop through a second time because it does funny things if you combine the `install.packages()` and `library()` steps in to one.
        if (!package %in% .packages()) { #`.packages()` returns a vector of all the loaded packages...
            suppressPackageStartupMessages (
                library ( package
                          , character.only = TRUE
                          , quietly = TRUE
                          , warn.conflicts = FALSE
                          , verbose = FALSE
                )
            )
        }
        if (!package %in% .packages()) {
            stop(paste0("Package '", package, "' was not loaded properly."))
        }
    }
    
    # Return
    return(TRUE)
    
}

#### Three string manipulation functions (LEFT,RIGHT,MID) ####
str_left <- function(string, num_chars) {
    
    # Input:
    # - 'string' is the text string you want to select from; must be an character type.
    # - 'num_chars' is the number of characters that you want to select; must be an atomic numeric type.
    
    # Output:
    # - A text string of length 'num_chars' that corresponds to the left most number of characters from the 'string' option.
    
    # Validations:
    stopifnot(is.character(string))
    stopifnot(is.numeric(num_chars))
    stopifnot(is.atomic(num_chars))
    
    # Do work
    return <- substr(string, 1, num_chars)
    
    # Return
    return(return)
    
}

str_mid <- function(string, start_num, num_chars) {
    
    # Input:
    # - 'string' is the text string you want to select from; must be an atopic string.
    # - 'start_num' is the starting position of the mid-text string you want to select from; must be an atomic numeric type.
    # - 'num_chars' is the number of characters that you want to select; must be an atomic numeric type.
    
    # Output:
    # - A text string of length 'num_chars' that corresponds to the characters from the 'start_num' starting position from the 'string' option.
    
    # Validations:
    stopifnot(is.character(string))
    stopifnot(is.numeric(start_num))
    stopifnot(is.atomic(start_num))
    stopifnot(is.numeric(num_chars))
    stopifnot(is.atomic(num_chars))
    
    # Do work
    return <- substr(string, start_num, start_num + num_chars - 1)
    
    # Return
    return(return)
    
}

str_right <- function(string, num_chars) {
    
    # Input:
    # - 'string' is the text string you want to select from; must be an character type.
    # - 'num_chars' is the number of characters that you want to select; must be an atomic numeric type.
    
    # Output:
    # - A text string of length 'num_chars' that corresponds to the right most number of characters from the 'string' option.
    
    # Validations:
    stopifnot(is.character(string))
    stopifnot(is.numeric(num_chars))
    stopifnot(is.atomic(num_chars))
    
    # Do work
    return <- substr(string, nchar(string) - (num_chars - 1), nchar(string))
    
    # Return
    return(return)
    
}

```


```{r LOAD Packages with Function, echo=FALSE, eval=TRUE, results="hide", warning=FALSE, message=FALSE}
LoadPackages(c("ggplot2", "dplyr", "magrittr", "tidyr", "knitr", "kableExtra", "scales"))
```


```{r SET Defaults, echo=FALSE, eval=TRUE}
# Set Default themes
theme_set(theme_bw())
theme_update(plot.title = element_text(hjust=0.5)
            ,plot.subtitle = element_text(hjust=0.5)
            )

# Set Default table and figure sizes
knitr::opts_chunk$set(rows.print=200, cols.print=30, fig.width=10, fig.height=7)

# Set Default rounding length
options(digits = 4)
options(scipen = 999)
```


<!-- Report -->

# Introduction {#Introduction}

The Net Promoter Score (NPS) is a trusted metric used by countless business to decide whether customers are Detractors or Promoters of the business. There is extensive resources online to justify the use of this metric, including on sites such as [qualtrics.com](https://www.qualtrics.com/experience-management/customer/net-promoter-score/), [wikipedia.org](https://en.wikipedia.org/wiki/Net_Promoter) and [netpromoter.com](https://www.netpromoter.com/know/).

The calculation of the NPS value is quite simple: $NPS = \%Promoters - \%Detractors$


There are two options that can be used to visualise the NPS Value:

1. First is in the true essence of the metric, which is a bar plot, the the NPS score displayed on the plot.
2. Second is to display a density plot for the data.

This Vignette is provides a helpful guide to visualise the NPS data for both these methods, using [`ggplot2`](https://ggplot2.tidyverse.org/) in the [`R`](https://www.r-project.org/) programming language.


# Set Up {#SetUp}

## Load the Packages {#LoadPackages}

To begin, the environment must be set up. The first step is to load the packages that will be used.

```{r LOAD Packages, echo=TRUE, eval=FALSE}
library(ggplot2)
library(dplyr)
library(magrittr)
library(tidyr)
```


## Generate the Data {#GenerateData}

The next step is to generate the NPS data. For this, the `sample()` function is used to generate $1,000$ values between $6$ and $10$. The first $20$ values are printed below for convenience.

Noting that this dummy data is generated by using a function. However, this data can be collected from any survey software, and fed in to this data pipeline at this point. The only prerequisite is that the data be a single vector of numbers that are all integers between $0$ and $10$, inclusive.

```{r GEN NPS Data, echo=TRUE, eval=TRUE}
set.seed(123)
NpsData <- sample(x=6:10, size=1000, replace=TRUE, prob=c(0.06, 0.05, 0.05, 0.42, 0.42))
NpsData %>% head(20) %>% print()
```


## Check the Data {#CheckData}

Next, to confirm that the data looks correct, the descriptive statistics are calculated for the generated data. For this, the `summarise_all()` function is used to calculate some key statistics.

```{r GEN Description of NPS Data, echo=TRUE, eval=FALSE}
NpsData %>%
    data.frame() %>% 
    summarise_all(list(Min=min, Max=max, Mean=mean, `Standard Deviation`=sd, Count=NROW)) %>%
    gather("Statistic", "Value") %>% 
    mutate_at("Value", round, 2)
```

```{r REVIEW Description of NPS Data, echo=FALSE, eval=TRUE}
NpsData %>%
    data.frame() %>% 
    summarise_all(list(Min=min, Max=max, Mean=mean, `Standard Deviation`=sd, Count=NROW)) %>%
    gather("Statistic", "Value") %>% 
    mutate_at("Value", round, 2) %>% 
    kable(align="l") %>% 
    kable_styling(bootstrap_options=c("striped","bordered","condensed")
                 ,full_width=FALSE
                 ,position="left"
                 ) %>% 
    (function(x){
        x %>% save_kable("Images/DescriptionOfNpsData.png")
        x %>% return()
    })
```


# Option One: Visualise Bar Plot {#BarPlot}

## Summarise NPS Data {#GenerateScore}

To calculate the NPS score, the following steps are performed on the data:

1. Coerce the data in to a `data.frame`;
1. Add a `Category` variable to determine the category of the score;
1. Count the number of scores in each category;
1. Calculate the percentage of the different categories;
1. Calculate the NPS score; and
1. Coerce it again in to a `data.frame` to add a new variable called `NPS`.

Once generated, the data is ready to be visualised.

```{r GEN Bar Summary Mean, echo=TRUE, eval=TRUE}
NpsScore <- NpsData %>% 
    data.frame(Score=.) %>% 
    mutate(Category="Promoters"
          ,Category=ifelse(Score<=8, "Passives", Category)
          ,Category=ifelse(Score<=6, "Detractors", Category)
          ,Category=factor(Category, levels=c("Promoters", "Passives", "Detractors"))
          ) %>% 
    count(Category, name="Count") %>% 
    mutate(Percentage=Count/sum(Count)) %>% 
    (function(x){
        Pro <- x %>% filter(Category=="Promoters") %>% select(Percentage) %>% pull()
        Det <- x %>% filter(Category=="Detractors") %>% select(Percentage) %>% pull()
        return((Pro-Det)*10)
    }) %>% 
    data.frame(Score=.) %>% 
    mutate(Name="NPS")
```

```{r REVIEW Bar Summary Mean, echo=FALSE, eval=TRUE}
NpsScore %>% 
    kable(align="l") %>% 
    kable_styling(bootstrap_options=c("striped","bordered","condensed")
                 ,full_width=FALSE
                 ,position="left"
                 ) %>% 
    (function(x){
        x %>% save_kable("Images/SummaryOfBarData.png")
        x %>% return()
    })
```


## Generate BarPlot Data Frame {#GenerateBar}

In order to properly visualise the NPS score, an empty data frame is generated, with one row being each of the possible scores. The reason for this is to allow for the Bar Plot to be adequately displayed. The way that this data is generated is by using the `seq()` function to create an ordered sequence of numbers from $0$ to $10$, incrementing by $1$ each time.

```{r GEN Bar Empty Data.Frame, echo=TRUE, eval=TRUE}
NpsFrame <- seq(from=0, to=10, by=1) %>% 
    data.frame(NPS=.) %>% 
    mutate(Name="NPS"
          ,Category="Promoters"
          ,Category=ifelse(NPS<9,"Passives",Category)
          ,Category=ifelse(NPS<7,"Detractors",Category)
          ,Category=factor(Category, levels=c("Promoters", "Passives", "Detractors"))
          ,NPS=factor(NPS, levels=0:10)
          )
```

```{r REVIEW Bar Empty Data.Frame, echo=FALSE, eval=TRUE}
NpsFrame %>% 
    kable(align="l") %>% 
    kable_styling(bootstrap_options=c("striped", "bordered", "condensed")
                 ,full_width=FALSE
                 ,position="left"
                 ) %>% 
    (function(x){
        x %>% save_kable("Images/BarEmptyDataFrame.png")
        x %>% return()
    })
```


## Join them all together {#JoinData}

Next, the NPS score and the NPS frame are joined together, so that the NPS score is replicated over each line. This is done by using the `left_join()` function, and using `Name` as the joining variable between the two frames.

```{r GEN Bar Joined Data.Frames, echo=TRUE, eval=TRUE}
FinalData <- left_join(x=NpsFrame
                      ,y=NpsScore
                      ,by="Name"
                      )
```

```{r REVIEW Bar Joined Data.Frames, echo=FALSE, eval=TRUE}
FinalData %>% 
    kable(align="l") %>% 
    kable_styling(bootstrap_options=c("striped","bordered","condensed")
                 ,full_width=FALSE
                 ,position="left"
                 ) %>% 
    (function(x){
        x %>% save_kable("Images/BarJoinedDataFrame.png")
        x %>% return()
    })
```

## Plot the final output {#PlotNps}

Finally, the result is plotted using the `ggplot()` function and the following layers: `geom_bar()`, `geom_point()`, and `geom_label()`.

The following steps were followed:

1. Pipe the `FinalData` data frame in to the `ggplot()` function, using the `Name` variable as the sole aesthetic variable.
1. Add a `geom_bar()` layer, using the `Category` variable to determine which colours to use to fill the column, then add a border around the categories using the colour 'DarkGrey', and give it a width of $0.5$ units.
1. Add a `geom_point()` layer, using the following arguments:

    1. '`data`' is created using an anonymous function. This is so that the data used by the `ggplot()` function can be manipulated, without using another external variable. The manipulation was effectively used to create a single NPS score which can be used in this layer.
    1. '`aes`' is the aesthetic used for the `y` axis; which in this instance is the NPS score. This is used to determine where on the plot the point should be placed.
    1. '`shape`' is a `plus` symbol, which is used to determine the exact location of the point, as convenient for the human eye to see.
    1. '`size`' is the size of the symbol, which in this instance is $25$ units.
    
1. Add a `geom_label()` layer, using the following arguments:

    1. '`data`' is again manipulated to determine the same value as used in `geom_point()`.
    1. '`stat`' is the statistic used to calculate the position of the label; which in this instance is the value `identity`, which effectively tells `ggplot` to use the own identity of the data, and not calculate any other statistic for the data.
    1. '`aes`' is used to determine that the `label` should be the value from the `Score` variable, and that it should be placed at the `Score` position on the `y` axis. Effectively, this aesthetic is used to decide _what_ the value of the label should be, and _where_ it should be place on the plot.
    1. '`size`' is used to determine the size of the label; which in this instance is $5$ units.
    
1. Determine how many breaks should be used, and the limits of the `y` axis, using the `scale_y_continuous()` layer.
1. Determine the colours that should be used in the three different Categories, using the `scale_fill_manual()` layer.
1. Hide the axis text for the `y` axis, using the `axis.text.y.left` argument of the `theme()` layer.
1. Flip the coordinates of the plot, so that it appears to be a bar from left to right, using the `coord_flip()` layer.
1. Label the axes, using the `labs()` layer, to ensure that the correct information is displayed in the correct positions.

```{r PLOT Bar Data, echo=TRUE, eval=TRUE, fig.width=15, fig.height=3, error=FALSE}
FinalData %>% 
    ggplot(aes(Name)) +
    geom_bar(aes(fill=Category), colour="darkgrey", width=0.5, alpha=0.5) +
    geom_point(data=function(x) {x <- x %>% select(Name, Score) %>% mutate(Score=round(Score,2)) %>% distinct()}
              ,stat="identity"
              ,aes(y=Score)
              ,shape="plus"
              ,size=25
              ) +
    geom_label(data=function(x) {x %>% select(Name, Score) %>% mutate(Score=round(Score,2)) %>% distinct}
              ,stat="identity"
              ,aes(y=Score, label=Score)
              ,size=5
              ) +
    scale_y_continuous(breaks=seq(0,10,1), limits=c(0,10), oob=squish) +
    scale_fill_manual(values=c("#66bd63", "#fdae61", "#d73027")) +
    theme(axis.text.y.left=element_blank()) +
    coord_flip() +
    labs(title="NPS Score"
        ,fill="Category"
        ,y="NPS Score"
        ,x="NPS"
        )
```

```{r SAVE Bar Plot, echo=FALSE, eval=TRUE}
ggsave(plot=last_plot()
      ,filename="Images/BarPlot.png"
      ,width=15
      ,height=3
      )
```


# Option Two: Visualise Density Plot {#DensityPlot}

## Generate DensityPlot Data Frame {#GenerateDensity}

In order to visualise the Density Plot, the data does not need to be summarised, but it is better to remain in its raw form. It does, however, need to undergo the following manipulations:

1. Coerce in to a `data.frame`; and
1. Add the `Category` variable.

```{r GEN Density Frame, echo=TRUE, eval=TRUE}
FinalFrame <- NpsData %>% 
    data.frame(Score=.) %>% 
    mutate(Category="Promoters"
          ,Category=ifelse(Score<=8, "Passives", Category)
          ,Category=ifelse(Score<=6, "Detractors", Category)
          ,Category=factor(Category, levels=c("Promoters", "Passives", "Detractors"))
          )
```

```{r REVIEW Density Frame, echo=FALSE, eval=TRUE}
FinalFrame %>% 
    head(10) %>% 
    kable(align="l") %>% 
    kable_styling(bootstrap_options=c("striped","bordered","condensed")
                 ,full_width=FALSE
                 ,position="left"
                 ) %>% 
    (function(x){
        x %>% save_kable("Images/DensityFinalDataFrame.png")
        x %>% return()
    })
```


## Visualise the DensityPlot data {#VisualiseDensity}

Once the Density data frame is generated, it can be visualised through `ggplot()`, using the following aesthetics: `geom_bar()` and `geom_density()`.

The following steps were used:

1. Pipe the `FinalData` data frame in to the `ggplot()` function, using the `Score` variable as the sole aesthetic.
1. Add a `geom_bar()` layer, using the `Category` variable to determine the colouers to use to fill the column, then add a border around the categories using the colour 'DarkGrey', and give it a transparency value of $0.3$.
1. Add a `geom_density()` layer, using an aesthetic `y` value to determine that this value should be a 'count' of the data, not a 'density' of the data, then give it a 'Blue' colour, and increase the size to $1$ unit.
1. Determine the colours that should be used for the three different Categories, using the `scale_fill_manual()` layer.
1. Determine the breaks and the limits of the `x` axis, using the `scale_x_continuous()` layer.
1. Remove the legend from the plot, using the `theme()` layer.
1. Add labels for the plot, using the `labs()` layer.

```{r PLOT Density Data, echo=TRUE, eval=TRUE, fig.width=15, fig.height=5}
FinalFrame %>% 
    ggplot(aes(Score)) +
    geom_bar(aes(fill=Category), colour="darkgrey", alpha=0.3) +
    geom_density(aes(y=..count..), colour="blue", adjust=3, size=1) +
    scale_fill_manual(values=c("#66bd63", "#fdae61", "#d73027")) +
    scale_x_continuous(breaks=seq(0,10,1), limits=c(-0.5,10.5)) +
    theme(legend.position="none") +
    labs(title="Density Plot of NPS"
        ,x="Score"
        ,y="Count"
        )
```

```{r SAVE Density Plot, echo=FALSE, eval=TRUE}
ggsave(plot=last_plot()
      ,filename="Images/DensityPlot.png"
      ,width=15
      ,height=5
      )
```


# Conclusion {#Conclusion}

As seen, the Net Promoter Score is a useful metric to see the percentage of customers who are Promoters, Passives or Detractors of the business. This metric can be visualised in a simple BarPlot, with a static value displayed on the chart, or it can be visualised as a DensityPlot, showing the proportion of customers in the different categories. Both of these methodologies are provided in this Vignette, with a step-by-step guide from data manipulation to plotting.


# Post Script {#PostScript}

**Publications**: This report is also published on the following sites:

1. RPubs: [RPubs/chrimaho/AusEnergyPrices](http://rpubs.com/chrimaho/PlottingNPS)
1. GitHub: [GitHub/chrimaho/PlottingNPS](https://github.com/chrimaho/PlottingNPS)
1. Medium: [Medium/chrimaho/AusEnergyPrices](https://medium.com/@chrimaho/2958b642a51f?source=friends_link&sk=382557ae6dd6227d004eb42a374fbb8f)

**Change Log:** This publication was modified on the following dates:

1. 